Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data

### 2021-01-31 paper/paper.tex
paper: Improve Aims and Method section

\documentclass[11pt,a4paper,USenglish,twocolumn]{article}
\usepackage{unir_paper}
\usepackage{subcaption}

%---------------------------
%título del trabajo y autor
%---------------------------
\title{Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data}
\author{C. D. Álvaro, J. Álvaro, C. A. Guzmán}
\date{24th of January, 2021}

@@

The developed method aims to improve current techniques of the state of the art.
We achieve improvements not only in terms of \emph{computational efficiency},
with lower computational requirements, but in \emph{usability}, reducing the number
of hyperparameters to get a good characterization of the analyzed clusters.

@@

Clusterix is defined as an unsupervised and non-parameterized tool, but it
critically depends on the selection of three radii.

\begin{itemize}
  \item \textbf{Inner radius} which defines a region with stars that belong to the OC and field stars.
  \item \textbf{Void region}. This region is excluded from the analysis since it is supposed
                to be a transition region with stars that could belong the OC.
  \item \textbf{External region} or field region.
                This is the outer region with stars that do not belong the open cluster.
\end{itemize}

@@

There are several clustering algorithms: KMeans, Mean-Shift Clustering, DBSCAN \cite{ester1996density}
among other. While each one behaves better according to the distribution of the objects
to clusterize, we have chosen KMeans by its simplicity and good results.

This algorithm works well and gives good results at first approach.
However, we need to set a large number of clusters to find the open cluster we are looking for.
This fact complicates the identification of the OC and many times the found cluster contains to many outliers.
For that reason, we have search for a KMeans refinement based on an artificial neural network.

The \emph{Unsupervised Deep Embedding for Clustering Analysis} model
or \emph{DEC} \cite{xie2016unsupervised} takes KMeans as its starting point,
but then, it trains an autoencoder to reduce the feature space
and pass this transformed data through a Clustering Layer which refines the previous selection.

@@

>>>>

\documentclass[11pt,a4paper,english,twocolumn]{article}
\usepackage{unir_paper}
\usepackage{subcaption}
\usepackage{tablefootnote}

%---------------------------
%título del trabajo y autor
%---------------------------
\title{Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data}
\author{CD. Álvaro, J. Álvaro, CAG. Guzmán}
\date{24th of January, 2021}

@@

The developed method aims to improve classic techniques of the state of the art.
We achieve improvements not only in terms of \emph{computational efficiency},
with lower computational requirements, but in \emph{usability}, reducing the number
of hyperparameters to get a good characterization of the analyzed clusters.

@@

Clusterix is defined as an unsupervised and non-parameterized tool, but it
critically depends on the selection of three radii in the studied region.
The inner radius should contain stars belonging or not to the cluster.
While the outermost radii define a ring which only contains stars that do
not belong to the cluster. In addition, it is also sensitive to other
parameters, such as the soften parameter \emph{h}, or other restrictions
related with the searched proper motions.

@@

There are several clustering algorithms: K-Means, Mean-Shift Clustering, DBSCAN \cite{ester1996density}
among other. While each one behaves better according to the distribution of the objects
to clusterize, we have chosen K-Means by its simplicity and good results.

This algorithm works well and gives good results at first approach.
However, we need to set a large number of clusters to find the open cluster we are looking for.
This fact complicates the identification of the OC and many times the found cluster contains to many outliers.
For that reason, we have search for a K-Means refinement based on an artificial neural network.

The \emph{Unsupervised Deep Embedding for Clustering Analysis} model
or \emph{DEC} \cite{xie2016unsupervised} takes K-Means as its starting point,
but then, it trains an autoencoder to reduce the feature space
and pass this transformed data through a Clustering Layer which refines the previous selection.

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{\columnwidth}
    \centering
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_pm_melotte_22.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_melotte_22.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_filtered_melotte_22.png}
    \end{subfigure}
  \end{subfigure}
  \medskip
  \begin{subfigure}{\columnwidth}
    \centering
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_hr_diagram_melotte_22.png}
      \caption{K-Means}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.png}
      \caption{DEC}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_filtered_melotte_22.png}
      \caption{DEC (filt.)}
    \end{subfigure}
  \end{subfigure}
  \caption{Evolution of Melotte 22 characterization}
  \label{fig:melotte_22_characterization_evolution}
\end{figure}

Figure \ref{fig:melotte_22_characterization_evolution} shows the evolution
of the characterization of Melotte 22. The first column corresponds to the
characterization obtained by applying K-Means alone to the dataset. The middle
column is the characterization made by the DEC model. And the column on the right
shows the characterization made by the DEC and having filtered the quantiles
lower than 0.10 and higher than 0.90. The top row shows the proper motion
configuration space while the bottom one shows the Hertzsprung–Russell diagrams.
